Untitled

### 2023-01-30 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
wip on discussion

(i) describe a framework comprised of:
  - a set of standards for implementing an accountable, reusable and updatable MOSHEM; and
  - a modelling toolkit (online repositories and software) for applying those standards to the authoring of ready4 modules, datasets and analyses.

>>>>

(i) describe a framework comprised of:
  - a set of standards for implementing an accountable, reusable and updatable MOSHEM; and
  - a modelling toolkit (online repositories and software) for applying those standards to the authoring of ready4 modules, datasets and analyses; and

### 2023-01-30 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd
wip on discussion

The strengths, limitations and issues raised by this work can be discussed in terms of relevance to future development of ready4, to health economists interested in undertaking other MOSHEMS and to funders of health economic research.

## Implications for the development of the ready4 MOSHEM
One of the main novel contributions of the framework we have developed is that it declares an explicit set of standards for the implementation and use of a MOSHEM. A practical benefit of specifying these standards is that it enabled us to develop a modelling toolkit to partially automate workflows for authoring, documenting and disseminating ready4 modules, datasets and analyses. The modelling toolkit's practical utility has been demonstrated through its application in creating all ready4 components used in a utility mapping study [@Hamilton2021.07.07.21260129] and the automation it facilitates makes the development and curation of a large and growing number of ready4 digital assets more tractable. The interoperability of the ready4 modules we have developed today is demonstrated in multiple example literate programs on the ready4 documentation website [@rfwn2022] and in the program used to implement the utility mapping analysis [@hamilton_matthew_2022_6129906].

The framework's standards also provide a mechanism to assess the extent to which ready4 meets explicit accountability, reusability and updatability goals.

@@

>>>>

The work we have described has implications for the future development of the ready4 model and highlights a number of issues for health economists and funders of health economic research.

## Implications for the development of the ready4 MOSHEM
The most direct implication of the development of the framework is that it makes it feasible for us to implement ready4 as a MOSHEM. The standards specified by the framework make it possible to partially automate workflows for developing and applying ready4 through use of the framework's modelling toolkit. The framework's standards also provide a mechanism to assess the extent to which ready4 meets explicit accountability, reusability and updatability goals. We have demonstrated the practical utility of the modelling toolkit by applying it to authoring, documenting and disseminating ready4 module libraries  [@hamilton_matthew_2022_6084467; @hamilton_matthew_2022_6084824; @hamilton_matthew_2022_6116701; @gao_caroline_2022_6130155; @matthew_p_hamilton_2021_5646669]], datasets [@DVN/HJXYKQ_2021; @DVN/DKDIB0_2021]; analyses [@hamilton_matthew_2022_6129906; @hamilton_matthew_2022_6416330; @hamilton_matthew_p_2022_6321821] and reporting templates [@hamilton_matthew_2022_6116385; @matthew_p_hamilton_2022_5976988] used in a utility mapping study [@Hamilton2021.07.07.21260129]. The automation (described in more detail on the project documentation website [@rfwn2022]) of significant sections of the workflows used in creating and sharing these artefacts has the potential to generate significant efficiencies as we extend ready4 to model help-seeking choice, spatial epidemiology, household populations and primary mental health services. The interoperability of the ready4 modules we have developed with the modelling toolkit is demonstrated in the program used to implement the utility mapping analysis [@hamilton_matthew_2022_6129906] and in multiple example literate programs published on the ready4 documentation website [@rfwn2022]. We have used some of these example programs in conjunction with toy data [@DVN/HJXYKQ_2021] to illustrate the potential for ready4 modules to facilitate study replication and transferability.

@@

<!-- Is it desirable for this to become default authoring tool or better to be replication tool. -->

### 2023-01-30 Markdown/Manuscript/Child_RMDs/MS_Imp_Example.Rmd
wip on discussion

We applied the framework to develop modules for developing utility mapping models, supplying those modules with data and implementing utility mapping analyses. The study for which these tools were developed and which applied them to developing utility mapping models for a clinical mental health sample of young people has been previously described [@Hamilton2021.07.07.21260129]. The outputs we authored using the framework's modelling toolkit include:

@@

When we assesses these study outputs against each framework standard, we conclude that 18 standards were wholly or mostly met (see Table \@ref(tab:checktb)). The two standards where the study outputs currently fall short are in reporting code coverage and including a user-interface. Both these items are scheduled to be addressed when we release production versions of the code libraries.

>>>>

We applied the framework to develop ready4 modules for developing utility mapping models, supplying those modules with data and implementing utility mapping analyses. These tools were created to develop utility mapping models for a clinical mental health sample of young people in a study that has been previously described [@Hamilton2021.07.07.21260129]. The framework's modelling toolkit created the following ready artefacts:

@@

When we assesses these study outputs against each framework standard, we conclude that 18 out of 20 standards were wholly or mostly met (see Table \@ref(tab:checktb)). The two standards where the study outputs currently fall short are in reporting code coverage and including a user-interface. Both these items are scheduled to be addressed when we release production versions of the code libraries.

### 2023-01-30 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
wip on discussion

Based on published guidance on computational modelling in health economics and other disciplines and our own experience, we have identified 20 standards that we believe are important for implementing a MOSHEM that is accountable (seven standards), reusable (nine standards) and updatable (four standards). The standards are specified in Table \@ref(tab:timelygls) and are discussed below.

@@

Storing model code and data in separate files and locations (as opposed to hard coding - embedding data into source code) can make it easier to apply models to different decision contexts and, when necessary, to selectively restrict access to data that are confidential, while disseminating all other model artefacts. Clear distinctions should be made between model modules (code that defines abstract data structures and the algorithms that can be applied to data described by these structures), model datasets (digital information such as parameter values, unit records, etc) and model analyses (code that links model datasets to model modules and specifies the algorithms to apply to the data associated with each module).

The coding practice of encapsulation [@8717448] can be used to help ensure that model modules continue to work as intended even when they are combined [@ready4oop2022]. In some cases, combining modules may mean new versions of modules have to be created to better account for interaction effects. The coding concept of inheritance [@8717448] can be used to efficiently achieve this objective as well as to facilitate selective editing of modules when transferring models to different decision contexts [@ready4oop2022]. Writing algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is recommended as good practice for scientific computing [@Wilson_2017]. Functions to implement model algorithms can be associated with module data structures (also known as a class) via a special type of function called a method. Model modules of a similar type or purpose can be efficiently distributed and documented by bundling them as code libraries. It is good practice to make available test or toy data to demonstrate the use of model algorithms [@Wilson_2017].

Statistical models are a common output of health economic evaluations, but they are often not reported in a format that enables others to confidently and reliably re-use them [@Kearns2013]. Open source approaches can help address this by disseminating code artefacts that enable easy and appropriate use of a statistical model to make predictions with new data. However, great care must be exercised when doing so if models are derived from data on human subjects as some software artefacts by default contain a copy of the source dataset. Such dataset copies must therefore be replaced (for example, with synthetic data) and the amended artefact's predictive performance then retested before any public release. Another way to make MOSHEMs easier to use is to develop simple user-interfaces for non-technical users. In the open source R language, such user-interfaces are typically developed with the Shiny package, for which a tutorial aimed at health economists is available [@SmithR2020].

@@

Continuous integration [@CI2017] tools can help verify that each code update passes multiple quality tests. Finally, using deprecation conventions that take an informative and staged approach to retiring old code and data reduces the risk that model revisions have unintended consequences on third party users.  The package lifeycle [@lifecycle2021] provides tools for R developers to consistently deprecate their code.

>>>>

Based on published guidance on computational modelling in health economics and other disciplines as well as our own experience in developing ready4, we have identified 20 standards that we believe are important for implementing a MOSHEM that is accountable (seven standards), reusable (nine standards) and updatable (four standards). The standards are specified in Table \@ref(tab:timelygls) and are discussed below.

@@

Storing model code and data in separate files and locations (as opposed to hard coding - embedding data into source code) can make it easier to apply models to different decision contexts and to selectively restrict access to data that are confidential, while disseminating all other model artefacts. Clear distinctions should be made between model modules (code that defines abstract data structures and the algorithms that can be applied to data described by these structures), model datasets (digital information such as parameter values, unit records, etc) and model analyses (code that links model datasets to model modules and specifies the algorithms to apply to data associated with each module).

The coding practice of encapsulation [@8717448] can be used to help ensure that model modules continue to work as intended when they are combined [@ready4oop2022]. In some cases, combining modules may mean new versions of modules have to be created to better account for interaction effects. The coding concept of inheritance [@8717448] can be used to efficiently achieve this objective as well as to facilitate selective editing of modules when transferring models to different decision contexts [@ready4oop2022]. Writing algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is recommended as good practice for scientific computing [@Wilson_2017]. Functions to implement model algorithms can be associated with module data structures (also known as a class) via a special type of function called a method. Model modules of a similar type or purpose can be efficiently distributed and documented by bundling them as code libraries. It is good practice to make available test or toy data to demonstrate the use of model algorithms [@Wilson_2017].

Statistical models are a common output of health economic evaluations, but they are often not reported in a format that enables others to confidently and reliably re-use them [@Kearns2013]. Open source approaches can help address this by disseminating code artefacts that enable easy and appropriate use of a statistical model to make predictions with new data. However, great care must be exercised when doing so if models are derived from data on human subjects as some software artefacts by default contain a copy of the source dataset. Such dataset copies must therefore be replaced (for example, with synthetic data) and the amended artefact's predictive performance then retested before any public release. Another way to make MOSHEMs easier to use is to develop simple user-interfaces for non-technical users.

@@

Continuous integration [@CI2017] tools can help verify that each code update passes multiple quality tests. Finally, using deprecation conventions that take an informative and staged approach to retiring old code and data reduces the risk that model revisions have unintended consequences on third party users.

### 2023-01-30 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
wip on discussion

- licensed (with GPL-3 by default);

@@

- quality assured (triggering continuous integration workflows, including any unit tests created by module library authors).

The ready4pack library is designed to integrate with our GitHub organisation and depends on the other two module authoring libraries. Methods from the ready4fun library [@ready4fun2022] are used to verify that functions for implementing module algorithms are written in a consistent house style and to use their standardised format to automatically generate basic documentation for each function. Methods from the ready4class [@ready4class2022] library are used to streamline and standardise the authoring of module data-structures. Like ready4fun, the ready4class library uses standardised naming conventions to automatically generate basic documentation for each module data-structure.

@@

When used in conjunction with repositories, the six R libraries provide strong support for implementing 17 framework standards. However, the software only weakly supports implementing the standards relating to safe dissemination of statistical models (R8) and user-interface development (R9) and does not yet provide any workflow tools to help implement the standard for deprecation conventions (U4). Standards not supported or weakly supported by our software can be met with existing developer tools in R.

>>>>

- licensed (using GNU GPL-3 [@GNUGPL2022] by default);

@@

- quality assured (each update triggers continuous integration workflows, including any unit tests created by module library authors).

The ready4pack library is designed to integrate with our GitHub organisation and depends on two other module authoring libraries. Methods from the ready4fun library [@ready4fun2022] are used to verify that functions for implementing module algorithms are written in a consistent house style. The standardised format is then used by ready4fun to automatically generate basic documentation for each function. Methods from the ready4class [@ready4class2022] library are used to streamline and standardise the authoring of module data-structures. Like ready4fun, the ready4class library uses standardised code implementation to automatically generate basic documentation for each module data-structure.

@@

When used in conjunction with repositories, the six R libraries provide strong support for implementing 17 out of 20 framework standards. Standards relating to safe dissemination of statistical models (R8), user-interface development (R9) and deprecation conventions (U4) are better met through using existing third party R libraries. Preparing statistical models for dissemination can be accomplished with standard R data management tools like the dplyr [@dplyr2022] and purrr [@dplyr2022] libraries. User-interfaces are typically developed with the shiny [@shiny2022] library, for which a tutorial aimed at health economists is available [@SmithR2020]. The library lifeycle [@lifecycle2021] provides tools for R developers to consistently deprecate their code.

### 2023-01-30 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
wip on discussion

To undertake this work we first needed to develop a framework for standardisation. Standardisation is an essential enabler of automation, collaboration, interoperability and transferability in modular and open source approaches. However, MOSHEMs remain rare and is unclear what standards are most appropriate to specify. A MOSHEMs is a computational modelling project, so standards developed for ready4 will need to be verified and implemented by software, some of which will need to be developed for this specific purpose.

>>>>

To undertake this work we first needed to develop a framework for standardisation. Standardisation is an essential enabler of automation, collaboration, interoperability and transferability in modular and open source approaches. However, we are not aware of any appropriate single source of guidance about what standards should be specified for MOSHEMs. To be deployed in computational models, the standards specified for ready4 will need to be implemented by software, some of which will need to be developed for this specific purpose.

### 2023-01-30 Markdown/Manuscript/Child_RMDs/MS_Tables_2.Rmd
wip on discussion

```{r cpkgs, tab.cap='Modelling toolkit R packages for developing and using MOSHEMs that meet framework standards', tab.id = 'cpkgs', echo = F, results = 'asis'}

>>>>

```{r cpkgs, tab.cap='Modelling toolkit R libraries for developing and using MOSHEMs that meet framework standards', tab.id = 'cpkgs', echo = F, results = 'asis'}

### 2023-01-30 Markdown/Manuscript/Child_RMDs/MS_Tables_3.Rmd
wip on discussion

                                           "All development code is stored on repos in a GitHub organisation [@ready4gh2022] and all archived releases are available in a Zenodo community [@ready4zen2022]. All non-confidential data is stored in repositories within a Harvard Dataverse collection [@ready4dv2022]. ",

@@

                                           "All code libraries include functions. The most complete list of functions for each library is available by clicking the 'Manual - Developer (PDF)' link on each package's documentation homepage (see item A2 above)",
                                           "All module data-structures and algorithms are distributed as code libraries.",

@@

                                           "All code is version controlled using Git and GitHub. All source code is available in a GitHub organisation [@ready4gh2022].",
                                           "Semantic versioning is used in all code. As no code library has yet been submitted to CRAN, only the development version extensions of each version number have been incremented to date.",
                                           "All six libraries use continuous integration (CI). CI results for each library can be viewed at a URL that concatenates 'https://github.com/ready4-dev/', the library name and '/actions' (e.g. https://github.com/ready4-dev/youthvars/actions)",
                                           "Retired code is deprecated using tools from the lifecycle R library (e.g. everything after “## DEPRECATED FNS” in https://github.com/ready4-dev/youthvars/blob/main/data-raw/fns/add.R ). Package vignettes and datasets are also deprecated e.g. https://ready4-dev.github.io/youthvars/articles/Replication_DS.html )"

>>>>

                                           "All development code is stored on repos in a GitHub organisation [@ready4gh2022] and all archived code releases are available in a Zenodo community [@ready4zen2022]. All non-confidential data are stored in repositories within a Harvard Dataverse collection [@ready4dv2022]. ",

@@

                                           "All code libraries include functions. The most complete list of functions for each library is available by clicking the 'Manual - Developer (PDF)' link on each package's documentation homepage (see item A2 above). The relationship between functions can be illustrated using an app: https://www.ready4-dev.com/docs/getting-started/software/libraries/dependencies/",
                                           "All module data-structures and algorithms are distributed as five code libraries [@hamilton_matthew_2022_6084467; @hamilton_matthew_2022_6084824; @hamilton_matthew_2022_6116701; @gao_caroline_2022_6130155; @matthew_p_hamilton_2021_5646669].",

@@

                                           "All code is version controlled using git [@git20XX] and GitHub [@github2007]. All source code is available in a GitHub organisation [@ready4gh2022].",
                                           "Semantic versioning is used in all code. As no code library has yet been submitted to CRAN [@CRAN2022], only the development version extensions of each version number have been incremented to date.",
                                           "All five libraries use continuous integration (CI). CI results for each library can be viewed at a URL that concatenates 'https://github.com/ready4-dev/', the library name and '/actions' (e.g. https://github.com/ready4-dev/youthvars/actions)",
                                           "Retired code is deprecated using tools from the lifecycle R library [lifecycle2021] (e.g. everything after “## DEPRECATED FNS” in https://github.com/ready4-dev/youthvars/blob/main/data-raw/fns/add.R ). Package vignettes and datasets are also deprecated e.g. https://ready4-dev.github.io/youthvars/articles/Replication_DS.html )"
